### 2018-03-21 misc.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Misc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{remove this section}

\subsection{The $\Delta r$ statistic from \cite{pigot2010shape}}

The $\Delta r$ statistic from \cite{pigot2010shape}:

\begin{equation}
  \Delta r=\frac{ln\left(\frac{L(T)}{L(\frac{1}{2}T)}\right)-ln\left(\frac{L(\frac{1}{2}T)}{L(0)}\right)}{ln\left(\frac{L(T)}{L(\frac{1}{2}T)}\right)+ln\left(\frac{L(\frac{1}{2}T)}{L(0)}\right)}
  \label{eq:delta_r}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Gamma statistic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Worse than nLTT \cite{janzen2015approximate}.

Worse than $\Delta r$, as gamma statistic depends on the size of the tree \cite{etienne2012prolonging}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{nLTT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

By \cite{janzen2015approximate}, outperforms gamma statistic and phylogenetic diversity.

In the R package 'nLTT'

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Normalized rooted branch score}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{Better distinguish NRBS and NRTS}

See \cite{heled2010bayesian}.

It is to estimate the distance between two trees:

First obtain the (non-normalized) rooted branch score:

\begin{equation}
  \left|T_{1},T_{2}\right|=\sqrt{\sum_{c\in T_{1}\cup T_{2}}\left(B\left(T_{1},c\right)-B\left(T_{2},c\right)\right)^{2}}
  \label{eq:root_branch_score}
\end{equation}

Where:

$T_{x}$ denotes tree $x$, 
$\left|T_{1},T_{2}\right|$ denotes the (absolute value of the) rooted branch score between tree T1 and T2
and $B(T,c)$ is 'the length of the branch connecting clade c to the tree if it is present in T, 0 otherwise'

Normalization:

\begin{equation}
  NRBS = \frac{1}{\left|T\right|}\frac{\sum_{i=0}^{N}\left|T,T_{i}\right|}{N}
  \label{eq:normalized_root_branch_score}
\end{equation}

Where:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Normalized root tree score}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

See \cite{heled2010bayesian}.

It is to estimate the distance between two trees:

First obtain the (non-normalized) rooted branch score:

\begin{equation}
  \left|T_{1},T_{2}\right|=\sqrt{\sum_{c\in T_{1}\cup T_{2}}\left(D\left(T_{1},c\right)-D\left(T_{2},c\right)\right)^{2}}
  \label{eq:root_branch_score_huh_again}
\end{equation}

Where $T_{x}$: tree x

$\left|T_{1},T_{2}\right|$: (absolute value of the) rooted branch score between tree T1 and T2

$D(T,c)=\intop_{b_{0}}^{b_{1}}\frac{1}{N_{b}}dt$: b is the branch connecting clade c to the tree if present in T, 0 otherwise.

To cite \cite{heled2010bayesian}: 'This is not a true metric, as branch length and population size are confounded'

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Phylogenetic diversity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Worse than nLTT \cite{janzen2015approximate}.