### 2018-03-29 20180326/appendix.tex
Start from scratch

\section{Appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other representations of the data in the Results section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_error_for_scr}
  \caption{
    Mean nLTT statistic distribution (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
  }
  \label{fig:error_for_scr}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
  \subfloat[]{
    %\includesvg[width=0.5\textwidth]{figure_error_sampling_representative}
  }
  \subfloat[]{
    %\includesvg[width=0.5\textwidth]{figure_error_sampling_representative_zoom}
  }

  \subfloat[]{
    %\includesvg[width=0.5\textwidth]{figure_error_sampling_representative_mean}
  }
  \subfloat[]{
    %\includesvg[width=0.5\textwidth]{figure_error_sampling_representative_mean_zoom}
  }

  \caption{
    The effect of sampling the youngest or oldest subspecies to represent a species
    on a sample of (subfigures a and b) or the mean (subfigures c and d) 
    nLTT statistic distributions (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    when sampling either the youngest or oldest incipient species as a representative (fill of boxplot),
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
    The subfigures a and c show the full distribution, where subfigures
    b and d show only the values below the horizontal dotted line.
    See also figure \ref{fig:error_expected_mean_dur_spec_sampling}
  }
  \label{fig:error_sampling_representative}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_error_low_high_ess}
  \caption{
    The effect of a low and high ESS
    on the mean nLTT statistic distribution (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    for either low or high ESS (fill of boxplot),
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
  }
  \label{fig:error_low_high_ess}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_error_alignment_length}
  \caption{
    The effect of alignment length
    on the nLTT statistic distribution (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    for different DNA alignment lengths (fill of boxplot),
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
    Bars at median,
    boxes ends at 25 and 75,
    whiskers at 5 and 95
  }
  \label{fig:error_alignment_length}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_error_alignment_length_mean}
  \caption{
    The effect of alignment length
    on the mean nLTT statistic distribution (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    for different DNA alignment lengths (fill of boxplot),
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
    Bars at median,
    boxes ends at 25 and 75,
    whiskers at 5 and 95
  }
  \label{fig:error_alignment_length_mean}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_error_alignment_length_median}
  \caption{
    The effect of alignment length
    on the median nLTT statistic distribution (each boxplot) 
    for different speciation completion rates (x axis boxplot)
    for different DNA alignment lengths (fill of boxplot),
    for different extinction rates (columns)
    and different speciation initiation rates (rows). 
    Bars at median,
    boxes ends at 25 and 75,
    whiskers at 5 and 95
  }
  \label{fig:error_alignment_length_median}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Diagnostics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_run_times}
  \caption{
    Distribution of simulation run times. All jobs finishing within 10 minutes are removed
  }
  \label{fig:run_times}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_exit_statuses}
  \caption{
    Exit statuses of simulations
  }
  \label{fig:exit_statuses}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_strees_different}
  \caption{
    Differences in the sampled species trees
  }
  \label{fig:strees_different}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_alignment_qualities}
  \caption{
    Informativeness of alignments. 
    OK: All taxa have a different DNA sequence.
    zero: At least one taxon pair has an identical DNA sequence.
    na: At least one taxon pair has an incalculable distance; 75% or more of the nucleotides differ.
  }
  \label{fig:alignment_qualities}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_exit_statuses_posteriors}
  \caption{
    Exit statuses of the posterior runs [Note: the be removed in final article]
  }
  \label{fig:exit_statuses_posteriors}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_ess_distribution}
  \caption{
    Distribution of ESSes for all estimated parameters. 
    The dotted line indicates an 'acceptable' ESS
  }
  \label{fig:ess_distribution}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_lowest_ess}
  \caption{
    Which parameter estimate has the lowest ESS?
  }
  \label{fig:lowest_ess}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  %\includesvg[width=\textwidth]{figure_acceptance_mcmc_operators}
  \caption{
    acceptance and rejection of MCMC operators. The horizontally
    dotted line is the target acceptance rate of 0.234 [BEAST book, page 89, quoting
    Gelman et al., 1996]. See table \ref{table:mcmc_operator_descriptions}
    for a description of these operators.
  }
  \label{fig:acceptance_mcmc_operators}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    Name & Description \\
    \hline
    \hline
    BirthRateScaler & Modifies speciation rate $lambda$ \\
    DeathRateScaler & Modifies extinction rate $mu$ \\
    SubtreeSlide & Changes the topology of the tree \\
    treeRootScaler & Changes $t_0$ \\
    treeScaler & Rescales entire tree \\
    \hline
  \end{tabular}
  \caption{
    MCMC operators.
  }
  \label{table:mcmc_operator_descriptions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%